\LARGE \bf Mean Estimation from Single-bit Measurements

### 2017-09-26 Writeups/onebit_distributed_V2.tex
Add files via upload

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference]{IEEEtran}      % Use this line for a4 paper
%\IEEEoverridecommandlockouts                              %

@@

\title{\LARGE \bf Mean Estimation from Adaptive Single-bit Measurements}
%
\author{
\IEEEauthorblockN{Alon Kipnis}
\IEEEauthorblockA{Department of Electrical Engineering \\
Stanford University\\
Stanford, CA\\}
\and
\IEEEauthorblockN{John C. Duchi}
\IEEEauthorblockA{Department of Electrical Engineering \\
and Department of Statistics \\
Stanford University\\
Stanford, CA\\}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the constraint: that the estimator can access only a single bit per sample from this distribution. We study the squared error in this estimation as a function of the number $n$ of samples and single-bit measurements in an adaptive setting. We show that if at each stage $n$, this single bit is a function of both the new sample and the previous $n-1$ acquired bits, then no estimator can attain asymptotic squared error smaller than $\pi/(2n)$ times the variance. Therefore, we conclude that a strict bit limitation increases the number of samples required for a prescribed accuracy by a factor of at least  $\pi/2$ compared to the unrestricted case. In addition, we provide an explicit estimator that attains this asymptotic error, showing that, rather surprisingly, only $\pi/2$ times more samples are required in order to attain performance equivalent to the unrestricted case. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}

Processing and estimating information from data collected at multiple physical location may be subject to communication constraints. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from data collected by the sensors is dictated not only by quality of observations and their numbers, but also by the constraint on the communication rates between the sensors. The question that we ask is to what extend a parametric estimation task is affected by such communication constraint, and what are the fundamental performance of estimation subject to these restrictions. \\

In this paper we provide an answer to this question in a particular simple setting: the estimation of the mean $\theta$ of a normal distribution with a known variance $\sigma^2$, under the constraint that given the $n$th sample $X_n$, only a single bit can be communicated to the estimator. 
Specifically, the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages $M_1,\ldots,M_{n-1}$, and delivers a single bit message $M_n$ to the estimator. The estimator provides an estimate $\widehat{\theta}_n$ based on $M^n = (M_1,\ldots,M_n)$.

 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. Since without bit constraint this risk is known to be $\sigma^2/n$, we are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting, defined as the ratio between their MSE to $\sigma^2/n$ as $n$ goes to infinity.
%\par
%Maybe explain here why the problem is interesting ? 
%setting (i), setting (ii) + SDM, setting (iii) + CEO.
%The turn to 1) main results and 2) other previous works. 
%
\par
Classical works on statistical estimation under communication constraint in information theory \cite{720540, zhang2013information, zhang1988estimation} implies that the ARE in setting (i) is $1$. Namely, asymptotically, there is no loss in performance in this setting due to the communication constraint. Indeed, under setting (i) the sample mean can be described by an encoder with an exponentially decaying error, leading to MSE of $\sigma^2/n + o(2^{-n})$. In this work we show that a similar result does not hold even in setting (ii): the relative efficiency of any adaptive estimator is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $\approx 1.57$ compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that almost surely attains this minimal relative efficiency. In other words, we show that the lower bound of $\pi/2$ on the asymptotic relative efficiency is tight, and that it is attained regardless of the particular realization of $\theta$ or the size of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether such efficiency is achievable (or otherwise, what is the minimal efficiency) remains open. \\

\subsection{Relate Works}
The parametric estimation problem we consider is closely related to many works in statistics and information theory. Statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation} includes as a special case the problem of estimating a single vector of parameters from multiple data-rate constrained observations. These inference problems were studied in the asymptotic limit as the number of observations goes to infinity. In this asymptotic setting the unconstrained inference performance is always attained when all samples are taken from the same distribution \cite[Sec. III]{720540}. Indeed, in an i.i.d setting the \emph{type} of the sample is a sufficient statistics for any estimation task, and the latter can be described using a number of codewords polynomial in $n$ regardless of the distribution of the samples. For this reason in \cite{han1987hypothesis} and \cite{zhang1988estimation} attention is given to inference problems involving multiple distributions observed at different locations. Setting (i) above can be seen as a special case of the above works where the communication is limited to a single-bit per terminal. In fact, with a full access to the sample prior to encoding, the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample \cite{gray1998quantization}.
%On the other hand, setting (iii) arises as the number of terminals grow with $n$ and their communication rates are limited to a single bit.

This last situation is also obtained by the multiterminal source coding setting, a.k.a. the CEO problem \cite{berger1996ceo} can be seen as a generalization of our setting (iii), where multiple draws of $\theta$ and rounds of $n$ bit communication is permitted. In particular, the minimal CEO distortion with $1$ bit per terminal provides a lower bound on the MSE risk.

In fact, with a full access to the sample as in setting (i), the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample \cite{gray1998quantization}. The sigma-delta modulatior (SDM) \cite{1092194} uses a single bit quantizer with a feedback loop, and therefore falls under setting (ii). The SDM with a constant input $\theta$ corrupted by a Gaussian noise was considered in \cite{53738}, where it was shown that the output of the modulator converges to the true constant input. Therefore, the SDM provides a consistent estimator for setting (ii). The rate of this convergence, however, was not analyzed and cannot be derived directly from the results of \cite{53738}. In fact, the results of this paper implies that the rate of convergence of a SDM to a constant input signal is at most $\sigma^2\pi/2$ over the number of feedback iterations. \par

%
Estimation problem subject to the one-bit quantization constraint were also considered in compressed sensing \cite{boufounos20081,baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}.

The rest of this paper is organized as follows: the main problem and notation are defined in Section~\ref{sec:problem}. In Section~\ref{sec:sequential} we present a lower bound on estimation from one-bit samples in the adaptive scheme, as well as an achievable scheme that attains this bound. Concluding remarks are given in Section~\ref{sec:conclusions}.

\section{Motivation and Related Works}
Standard considerations implies that the optimal encoding in setting (i) is reduced to an optimal $n$ bits description of the  sample mean. The standard technique to encode an unknown random quantity using $n$ bits is attained by an optimal scalar quantizer. However, the design of this quantizer depends on the distribution of its input \cite{gray1998quantization}, which is the goal of the estimation. Therefore, although the ARE in this setting is shown to be one \cite[Sec. III]{720540}, we see that a non-trivial challenge arise even in this case. 
In setting (ii), the above challenge of optimal quantizer design with imprecise input distribution occurs upon the the arrival of each new sample $X_n$.

%sigma delta modulator

%implies the existence of a consistent estimator in this case

% our result provides a lower bound for the accuracy of a sigma-delta modulator

Setting (ii) and (iii) were considered in 
\cite{zhang2013information} under the more general assumption of $m$ machines each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are lower bounds on the estimation error as a function of the total number of bits $R$ with which each machine can communicate. Our settings corresponds to the case $m=n$ and $R=1$, and our result is a fundamental lower bound and a matching achievable scheme under this simplified setting.

 We also note that the counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.

 The optimal design  requires the distribution of the input, which is the goal of estimation and is known appriori.

 standard technique to encode an unknown random quantity using $n$ bits is attained by an optimal scalar quantizer,  by designing a scaler quantizer Unlike standard quantization task of encoding an unn

\subsection{Intuition from Remote Multiterminal Source Coding}
A first clue for this rather surprising result can be seen drawing the connection between our parametric estimation problem to the remote multiterminal source coding problem of \cite{berger1996ceo}, also known as the CEO problem. The CEO setting considers $n$ encoders, each has access to a corrupted version of a source signal. The $i$th encoder observes $k$ noisy source symbols and transmit $R_i k$ bits to a central estimator. Therefore, assuming that $\theta$ is drawn randomly from a prior $\pi(\theta)$, the distributed estimation setting (iii) corresponds to the CEO setting with $k=1$, where the $i$th agent observes 
\begin{equation}
\label{eq:Gaussian_channel}
X_i = \theta + Z_i,
\end{equation}
and has $R_i=1$ bits to transmit this observation. As a result, the quadratic distortion in the optimal source coding scheme for the CEO with $n$ terminals at rates $R_1 = \ldots = R_n = 1$ and Gaussian observation noise of variance $\sigma^2$ provides a lower bound on the MSE distortion in estimating $\theta$. In fact, the difference between this lower bound and the actual MSE in our setting indicates the importance of coding over blocks consisting of $k$ independent realizations of $\theta$ versus $k=1$ in our setting. \par
A closed-form expression for CEO under quadratic distortion is known only for the case where the sequence to described by the encoders is sampled from a Gaussian distribution \cite{prabhakaran2004rate}. By using the characterization of the minimal CEO distortion as the number of terminals goes to infinity, we conclude the following:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\pi(\theta)$ is the normal distribution. Then any estimator $\widehat{\theta}_n$ of $\theta$ in the distributed setting satisfies
\[
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4\sigma^2}{3} + o(1).
\]
\end{prop}

\begin{proof}
We consider the expression \cite[Eq. 10]{chen2004upper} that provides the minimal distortion $D^\star$ in the CEO with $L$ observers and under a total sum-rate $R_\Sigma = R_1 + \ldots R_L$:
\begin{equation} \label{eq:ceo_optimal_sumrate}
R_{\Sigma} = \frac{1}{2} \log^+ \left[ \frac{\sigma_\theta^2}{D^\star} \left( \frac{D^\star L}{ D^\star L - \sigma^2 + D^\star \sigma^2 / \sigma_\theta^2 }\right)^L  \right].
\end{equation}
Assuming $R_\Sigma = n$ and $L=n$ we get
\begin{equation} \label{eq:ceo_optimal_sumrate}
n = \frac{1}{2} \log \left[ \frac{\sigma_\theta^2}{D^\star} \left(\frac{ D^\star n }{D^\star n - \sigma^2 + D^\star \sigma^2/\sigma_\theta^2 }  \right)^n  \right].
\end{equation}
The value of $D^\star$ that satisfies the equation above describes the MSE in the quadratic Gaussian CEO setting under an optimal allocation of the sum-rate $R_\Sigma = n$ among the $n$ encoders. Therefore, $D^\star$ provides a lower bound to the CEO distortion with $R_1=\ldots,R_n = 1$ and hence a lower bound to the minimal MSE in estimating $\theta$ in the distributed setting. By considering $D^\star$ in \eqref{eq:ceo_optimal_sumrate} as $n\rightarrow \infty$ we conclude that 
\[
D^\star = \frac{ \frac{4\sigma^2}{3} }{n + \frac{4 \sigma^2}{3 \sigma_\theta^2} } + O(e^{-n}) =  \frac{4\sigma^2}{3n} + o(n^{-1}). 
\]
\end{proof}
We note that although the CEO lower bound was derived by the MSE attainable assuming optimal allocation a total of $n$ bits among the encoders, the same expression is from the upper bound to the CEO derived in \cite[Prop. 5.2]{KipnisRini2017}, that assumes $R_1=\ldots,R_n = 1$ rather than optimal allocation of $n$ bits among all encoders. %\[
%D_{CEO} / \sigma_\theta^2 \leq \left( 1 + \sum_{i=1}^n \frac{\sigma_\theta^2}{\sigma^2} \frac{1 - 2^{-2R_i}} {1+ \frac{\sigma_\theta^2}{\sigma^2}  2^{-2 R_i}  } \right)^{-1}
%\]
%where $\sigma_\theta^2$ is the variance of the distribution $\pi(\theta)$. Using $R_1=\ldots=R_n=1$ we get
%\[
%D_{CEO} \leq  \left( \frac{1}{\sigma_\theta^2} +  \frac{3n}{4\sigma^2 + \sigma_\theta^2} \right)^{-1}   =
%\frac{4 \sigma^2}{3n} +  \frac{\sigma_\theta^2}{3n} + o(n^{-1}). 
%\]

Statistical inference problems under communication constraints in the nature of our setting were considered in information theory in 
\cite{han1987hypothesis, zhang1988estimation, 720540}. These works consider a multi-terminal source coding setting in which the $i$th terminal observes $n$ samples from the distribution and is allotted $nR_i$ bits to communicate its estimate. The main focus of these works is the difference between inference with communication constraint and the unconstrained vanilla statistical estimation setting, as the number of samples $n$ goes to infinity subject to a total finite rate constraint. Our setting (i) can be seen as a special case of this setting where with a single terminal and $R_1 = 1$.% whereas our case (iii) is obtained from this setting by taking $n=1$, $R_i=1$, and number of terminals grow with $n$. 
However, as explained in \cite[Sec. III]{720540}, a single terminal in this setting always lead to the unconstrained inference performance. Indeed, when all samples are taken from the same distribution, the \emph{type} of the sample \cite{csiszar1998method} is a sufficient statistics for any inference task, and the latter can be described using a number of codewords polynomial in $n$ regardless of the distribution of the samples. For this reason, attention is given in these works to inference problems involving multiple distributions observed at different locations and hence the results there do not apply to our setting. In fact, with a full access to the sample as in setting (i), the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample. Finally, we note that setting (ii) includes as a special case the sigma-delta modulation (SDM) analog-to-digital conversion scheme with a constant input $\theta$ corrupted by Gaussian noise $Z_i$, as was considered in \cite{53738}. While it was shown there that the output of the modulator converges to the true constant input, the rate of this convergence was not analyzed and cannot be derived directly from the results of \cite{53738}. In fact, a corollary from the results in this paper we conclude that the rate of convergence of a SDM to a constant input signal is at most $\sigma^2\pi/2$ over the number of feedback iterations. Finally, the multiterminal source coding setting denoted as the CEO problem \cite{berger1996ceo} can be seen as a generalization of our setting (iii), where multiple draws of $\theta$ and rounds of $n$ bit communication is permitted. In particular, the minimal CEO distortion with $1$ bit per terminal provides a lower bound on the MSE risk.  \par

%Why it is surprising that the estimator is independent of $\theta$, $\Theta$ or $\pi(\theta)$.

\subsection{Related Works}
\subsubsection*{Statistical inference under communication constraints}
The problem of statistical inference under communication constraint has a long and rich history in statistics and information theory. Statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation} includes as a special case the problem of estimating a single vector of parameters from multiple data-rate constrained observations. These inference problems were studied in the asymptotic limit as the number of observations goes to infinity. The main question asked there is the difference between inference with communication constraint and the unconstrained vanilla statistical estimation setting. In this asymptotic setting the unconstrained inference performance is always attained when all samples are taken from the same distribution \cite[Sec. III]{720540}. Indeed, in an i.i.d setting the \emph{type} of the sample is a sufficient statistics for any estimation task, and the latter can be described using a number of codewords polynomial in $n$ regardless of the distribution of the samples. For this reason in \cite{han1987hypothesis} and \cite{zhang1988estimation} attention is given to inference problems involving multiple distributions observed at different locations. Setting (i) above can be seen as a special case of \cite{han1987hypothesis, zhang1988estimation, 720540} where the communication is limited to a single-bit per terminal.

@@

\subsubsection*{Relation to channel coding}
The problem we consider is the estimation of the parameter $\theta$ by observing 
\begin{equation}
\label{eq:channel}
X_i = \theta + Z_i, \quad i=1,\ldots,n,
\end{equation}
where the $Z_i$s are independent, centered normal with variance $\sigma^2$. The problem of estimating $\theta$ from the $X_i$s can be seen as the attempt to transmit a real number $\theta$ over an additive white Gaussian noise (AWGN) channel. In particular, setting (ii) can be seen as the feedback version of this transmission, which is closely related to \cite{horstein1963sequential, 1053879}. However, in our setting $\theta$ is either fixed or drawn once from a prior distribution, so the communication in \eqref{eq:channel} is limited to a repetition code. Moreover, in setting (ii) we have the additional constraint that the message is a causal function of the previous messages, where in case (iii) it is completely memoryless. Works in information theory which consider such setting includes AWGN channel with 1-bit ADC at the receiver \cite{DBLP:journals/corr/VarastehSG16}. \par

\subsubsection*{Relation to source coding}
With a full access to the sample as in setting (i), the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample. Setting (ii) includes as a special case the sigma-delta modulation (SDM) analog-to-digital conversion scheme with a constant input $\theta$ corrupted by Gaussian noise $Z_i$, as was considered in \cite{53738}. While it was shown there that the output of the modulator converges to the true constant input, the rate of this convergence was not analyzed and cannot be derived directly from the results of \cite{53738}. As a corollary from the results in this paper we conclude that the rate of convergence of a SDM to a constant input signal is at most $\sigma^2\pi/2$ over the number of feedback iterations. Finally, the remote multiterminal source coding setting of \cite{berger1996ceo} corresponds to the case of $n$ rate-constrained encoders, each observing a noisy version of an information source. The difference between this setting and ours is that in ours the parameter of interest is not an information source. By assuming a prior distribution on this parameter, the CEO provides a lower bound on the estimation error in the fully distributed setting (iii). This lower bound can be attained if we were to consider the average error in multiple independent realizations of our problem rather than a single realization as we do here. \\

\begin{comment}
\subsection{Sigma-Delta Encoding}

What is the error in estimating the mean of stationary ergodic process ? 
}
Since the SDM is a special case of the sequential scheme, we conclude that the MSE of any SDM with a noisy DC signal is bounded from below by $n^{-1} \sigma^2 \pi/2 $.
\end{comment}

The rest of this paper is organized as follows: the main problem and notation are defined in Section~\ref{sec:problem}. Our main results are presented and discussed in Section~\ref{sec:sequential}. Concluding remarks are given in Section~\ref{sec:conclusions}.

@@

@@

\node[above of = source, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

@@

Let $X_i$, $i=1,\ldots,n$, be $n$ independent samples from the normal distribution $\mathcal N(\theta,\sigma^2)$ with mean $\theta$ and variance $\sigma^2$. We assume that the mean $\theta$ is drawn once from a prior distribution $\pi(\theta)$ on the parameter space $\Theta$, which is a closed interval of the real line. We moreover assume that $\pi(\theta)$ is absolutely continuous with respect to the Lebesgue measure with density $\pi(d\theta)$. The problem we consider is the estimation of the parameter $\theta$ under the following constraints on the communication between the samples $X_1,\ldots,X_n$ and a centralized estimator: 
\begin{itemize}
\item[(i)] The estimator at time $n$ is only a function of the $n$ messages $M^n = \left(M_1,\ldots,M_n \right)$.
\item[(ii)] For each $i=1,\ldots,n$, the $i$th message $M_i$ is a function of the sample $X_i$ and the $i-1$ previous messages $M^{i-1}$.
\item[(iii)] The $i$th message $M_i$ takes only two possible values, say $1$ and $-1$. 
\end{itemize}
In other words,  the $n$ messages $M^n$ are the only information on the sample $X^n = \left(X_1,\ldots,X_n\right)$ available to the estimator. The $i$th message is defined by a function from the real line to $\{-1,1\}$, measurable with respect to the sigma algebra generated by $M^{i-1}$ and $X_i$.  The estimator, upon observing $M^n$, produces an estimate $\widehat{\theta}_n(M^n)$ of $\theta$. A system describing the above scheme is illustrated in Fig.~\ref{fig:sequential}. \\

In this work we are concerned with the mean squared error (MSE) risk:
\begin{equation}
\label{eq:error_def}
\epsilon_n \triangleq \mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X^n$ and the prior distribution $\pi(\theta)$. \\
%It is well known that minmax estimation error 
%\begin{equation}
%E_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left(\widehat{\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $\epsilon_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}.

>>>>

%\documentclass[letterpaper]{IEEEtran}
\documentclass[letterpaper, conference]{IEEEtran}      % Use this line for a4 paper

%LatexiDiff ignore tikz:
% $: latexdiff -c "PICTUREENV=(?:picture|tikzpicture|DIFnomarkup)[\w\d*@]*" old.tex new.tex > diff.tex

%\IEEEoverridecommandlockouts                              %

@@

\title{\LARGE \bf Mean Estimation from Single-bit Measurements}
%
%\author{
%\IEEEauthorblockN{Alon Kipnis}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{John C. Duchi}
%\IEEEauthorblockA{Department of Statistics \\
%and Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%\and
%\IEEEauthorblockN{Andrea J. Goldsmith}
%\IEEEauthorblockA{Department of Electrical Engineering \\
%Stanford University\\
%Stanford, CA\\}
%}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the constraint that only a single bit per sample from this distribution is available to the estimator. We study the mean squared error risk in this estimation as a function of the number of samples, and hence number of bits, from this distribution. We consider an adaptive scheme in which each bit is a function of the current sample and the previously observed bits, as well as a fully distributed setting in which each bit is only a function of the current sample. For the adaptive scheme, we show that the optimal relative efficiency, compared to the standard mean estimator without bit limitation, is $\pi/2$. Namely, the number of samples required to attain the mean to a prescribed accuracy is only $\pi/2$ times larger due to the bit constraint. For the distributed scheme we consider the setting where each bit is obtained by comparing the current sample to a prescribed value. For this setting, we show that the expected asymptotic efficiency converges to $\pi/2$ provided that the variance of the distribution is high compared to the apriori uncertainty about the mean. Our results indicate that, rather surprisingly, parametric estimation from the coarsest form of quantization without significant penalty in the number of measurements is possible. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:Intro}
Processing information on multiple physical locations may impose communication constraints on the amount of information that can be shared among different units of the system. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
%For example, this constraint occurs in large-scale sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. 
In this situation, the ability to estimate a particular parameter is dictated not only by quality of observations but also by the constraint on the communication rate of the sensor, which may be severely restricted due to power, bandwidth, or size of data. The question that we ask is to what extent a parametric estimation task is affected by such communication constraints, and what are the fundamental performance of estimation subject to these restrictions. \\

In this paper we provide a precise answer to this question in a limited setting: the estimation of the mean $\theta$ of a normal distribution with a known variance $\sigma^2$, under the constraint that only a single bit can be communicated to the estimator on each sample from this distribution. As it turns out, the ability to share information before communicating it dramatically affects the performance in recovering $\theta$. We distinguish in particular among three different settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all the $n$ encoders confer and produce a single $n$ bit message. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $i$th encoder observes the $i$th sample and the $i-1$ previously encoded observations.
 \item[(iii)] \emph{Distributed} encoding: the output of the $i$th encoder is only a function of the $i$th observation.
 \end{itemize}
As far as information sharing is concerned, it is evident that settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We are concerned with the mean squared error (MSE) risk in estimating $\theta$ as the number of observations $n$ goes to infinity. In particular, we are interested in the \emph{relative efficiency} of estimators in settings (i)-(iii), defined as the ratio between their MSE to $\sigma^2/n$. The latter is the MSE attained by the sample mean, which is the minimal risk estimator when no constraints on the communication are imposed. \par
%
It is well known that the relative efficiency in setting (i) is $1$ \cite{720540, zhang2013information, zhang1988estimation}, i.e., asymptotically, there is no loss in performance due to the communication constraint in this setting. Indeed, under setting (i) the sample mean can be described with an error exponentially decaying in $n$, leading to MSE of $\sigma^2/n + o(n)$. In this work we show that a similar result does not hold even in setting (ii): the relative efficiency of any adaptive estimator is at least $\pi/2$. Namely, a single bit per sample communication constraint in the adaptive setting incurs a minimal penalty of at least $\pi/2 \approx 1.57$ compared to an unconstrained estimator, or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator which asymptotically attains this minimal relative efficiency, i.e., we show that this lower bound is tight. Moreover, we show that even in the distributed setting (iii), a relative efficiency of $\pi/2$ is possible as long as $\sigma^2$ is large compared to the variance of the prior distribution on $\theta$. These results are quite surprising since they show that, even under the coarsest form of quantization, an estimator based on $ \approx 1.59 n$ measurements achieves accuracy in MSE or confidence interval comparable to an estimator that uses $n$ samples with infinite bit-precision.  \par
%
%One clue for this surprising results is given in a famous problem in information theory denoted as the CEO problem, first presented by Berger et. al. in \cite{berger1996ceo}. In this problem, a firms CEO is required to estimate a signal  from information sent to it by multiple agents, where the $i$th agents is given a noisy observation of this signal and has a limited number of bits, say $R_i$, to describe its observation to the CEO.

\subsection*{Related Works}
%\subsubsection*{Statistical inference under communication constraints}
Statistical inference problems under communication constraints were considered in information theory in 
\cite{han1987hypothesis, zhang1988estimation, 720540}. These works consider a multi-terminal source coding setting in which the $i$th terminal observes $n$ samples from the distribution and is allotted $nR_i$ bits to communicate its estimate. The main focus of these works is the difference between inference with communication constraint and the unconstrained vanilla statistical estimation setting, as the number of samples $n$ goes to infinity subject to a total finite rate constraint. Our setting (i) can be seen as a special case of this setting where with a single terminal and $R_1 = 1$.% whereas our case (iii) is obtained from this setting by taking $n=1$, $R_i=1$, and number of terminals grow with $n$. 
However, as explained in \cite[Sec. III]{720540}, a single terminal in this setting always lead to the unconstrained inference performance. Indeed, when all samples are taken from the same distribution, the \emph{type} of the sample \cite{csiszar1998method} is a sufficient statistics for any inference task, and the latter can be described using a number of codewords polynomial in $n$ regardless of the distribution of the samples. For this reason, attention is given in these works to inference problems involving multiple distributions observed at different locations and hence the results there do not apply to our setting. In fact, with a full access to the sample as in setting (i), the problem of encoding and estimating $\theta$ is reduced to the MSE attained by a scalar quantizer adjusted to the sufficient statistics of the sample. Finally, we note that setting (ii) includes as a special case the sigma-delta modulation (SDM) analog-to-digital conversion scheme with a constant input $\theta$ corrupted by Gaussian noise $Z_i$, as was considered in \cite{53738}. While it was shown there that the output of the modulator converges to the true constant input, the rate of this convergence was not analyzed and cannot be derived directly from the results of \cite{53738}. In fact, a corollary from the results in this paper we conclude that the rate of convergence of a SDM to a constant input signal is at most $\sigma^2\pi/2$ over the number of feedback iterations. Finally, the multiterminal source coding setting denoted as the CEO problem \cite{berger1996ceo} can be seen as a generalization of our setting (iii), where multiple draws of $\theta$ and rounds of $n$ bit communication is permitted. In particular, the minimal CEO distortion with $1$ bit per terminal provides a lower bound on the MSE risk.  \par
%The difference between this setting and ours is that in ours the parameter of interest is not an information source. By assuming a prior distribution on this parameter, the CEO provides a lower bound on the estimation error in the fully distributed setting (iii). This lower bound can be attained if we were to consider the average error in multiple independent realizations of our problem rather than a single realization as we do here. \\
%
%While our focus in this paper is on setting (ii), we note that unlike the above works, we are also interested in the optimal estimation performance for a finite number of observations and not only in their asymptotic behavior. 
%\subsubsection*{Relation to channel coding}
More related to our setting is the work \cite{zhang2013information} that considers the minimax MSE risk in estimating a vector of parameters from $m$ machines, where each machine has access to $n/m$ samples a limited number of bits to describe these samples. The main results of this work are lower bounds on the estimation error as a function of the total number of bits can be sent by each machine. In the terminology of \cite{zhang2013information}, our settings (ii) and (iii) fall under the category of \emph{interactive}, and  \emph{independent} communication protocols, respectively. %In both cases we use $m=n$ (single sample per machine) and communication constraint of a single bit per machine.
However, unlike in \cite{zhang2013information}, we focus on achievable schemes and tight bounds in $n$ for the asymptotic MSE risk, rather than its scaling rate. \\

The rest of this paper is organized as follows: the main problem is defined in Section~\ref{sec:problem}. Our main results are presented and discussed in Sections~ \ref{sec:sequential} and \ref{sec:distributed}. Concluding remarks are given in Section~\ref{sec:conclusions}.

%We also note that the counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.

@@

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]

\node[node distance = 1.5cm] at (0,0) (source2) {$X_1,\ldots,X_n$};
\node[int1, right of = source2, node distance = 2cm] (enc2) {Enc};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[above of = source2, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M \in \left\{-1,1\right\}^n$} (est);

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:centralized} Centralized encoding using one bit per sample on average.}
\end{figure}

@@

\node[below of = source2, node distance = 0.5cm] {$\vdots$};
\node[above of = source3, node distance = 0.6cm] {$\vdots$};
\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

@@

\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};
\node[above of = source3, node distance = 0.6cm] {$\vdots$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);

\draw[->,line width = 2pt] (enc3) -| (est);

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:distributed} Distributed single-bit encoding: the single-bit message produced by each encoder is only a function of its private sample $X_i$.}
\end{figure}
